%% This is an example first chapter.  You should put chapter/appendix that you
%% write into a separate file, and add a line \include{yourfilename} to
%% main.tex, where `yourfilename.tex' is the name of the chapter/appendix file.
%% You can process specific files by typing their names in at the
%% \files=
%% prompt when you run the file main.tex through LaTeX.
\section{Graphs and intersections}
\label{sec:graphs}


\subsection{Graphs}

A graph $G$ is defined as $G = (V,E)$, where $V$ is the set of vertices and $E$ the set of
edges, where $E \subseteq \binom{V}{2}$. The vertices $v,w \in V$ such that $e = vw \in E$
links are called the \textit{endpoints} of $e$.

\begin{defn}
  An embedding of a graph $G$ into a surface $\Sigma$ is a mapping of $G$ in
  $\Sigma$ where the vertices correspond to distinct points and the edges
  correspond to simple arcs connecting the images of their endpoints.
  \cite{goyalGraphEmbeddingTechniques2017}.
\end{defn}

A graph $G$ is planar if there is an embedding of this graph that does not have
any crossing between the edges.

\begin{defn}
  Let $G = (V,E)$ and $S \subset V$, an induced subgraph is a graph $H$ of $G$ whose
  vertex set is $S$ and its edge set $F = \{vw : v,w \in S, vw \in E\}$.
\end{defn}

\begin{defn}
  Let $G = (V,E)$ its complement graph $\overline{G}$ is the graph such that its edge set
  is defined as: $\{vw: v,w\in V, vw\notin E\}$.
\end{defn}

\begin{defn}
  $H$ is called a \textit{minor} of $G$ if $H$ can be constructed by deleting edges and vertices,
  or contracting edges.
\end{defn}

\begin{theorem}[Kuratowski]
  \label{theo:kura}
  A graph $G$ is planar if and only if it does not contain $K_5$ or $K_{3,3}$ as a minor or
  a induced subgraph.
\end{theorem}

\begin{defn}
  A path $P_n$ in a graph $G$ is a sequence of vertices $v_1v_2v_3\dots v_n$ such
  that $v_iv_{i+1} \in E$.
\end{defn}

\begin{defn}
  A cycle $C_n$ in a graph $G = (V,E)$ is a path $v_1\dots v_n$ such that $v_1 = v_n$.
\end{defn}

\begin{defn}
  A chord of a cycle $C_n$ with $n \geq 4 $ is an edge that connects two non consecutive vertices
  of $C_n$.
\end{defn}

\begin{defn}
  A triangular chord of a cycle is a chord that will create a new triangle ($C_3$).
\end{defn}


\begin{defn}
  A graph $G = (V,E)$ is complete if every pair of distinct $v_1,v_2 \in V$ are
  adjacent. This is denoted $K_n$ with $n$ the size of the graph. If $G$ is an
  induced graph of $H$ then $G$ is a clique of $H$.
\end{defn}

\begin{defn}
  A graph $G$ is bipartite if there exist two disjoint subsets $A,B \subset V$ such
  that $A\cup B = V$ and each edge $e\in E$ has an endpoint on $A$ and the
  another on $B$.
\end{defn}

\begin{defn}
  A bipartite graph $G$ with bipartitions $A$ and $B$ is complete bipartite if
  every pair of vertices $v\in A, w\in B$ are adjacent. It is denoted as $K_{n,m}$,
  being $n$ and $m$ the size of each bipartition.
\end{defn}

Some graphs can be characterized with properties. A property of a graph is a property that
is preserved under all its isomophisms. These properties are called \textit{hereditary} if they
are also preserved under all its induced subgraphs; they are called \textit{minor-hereditary} if they are
also preserved under its minors (e.g. Kuratowski's planar graph characterization [\ref{theo:kura}]).

\begin{defn}
  An forbidden induced subgraph (minor) of a graph class $X$ is a graph such that if it is the induced
  subgraph (minor) of a graph $G$, we know that $G \notin X$.
\end{defn}

The coloration of a graph is a color assignment to each vertex such that the color
of the two endpoints of every edge of the graph is different.

\begin{defn}
  The chromatic number of a graph $\chi(G)$ is the smallest number of colors needed to
  have an acceptable coloration of $G$.
\end{defn}

\begin{defn}
  The clique number of a graph $\omega(G)$ is the size of the biggest clique of $G$. We
  can observe that for every graph: $\chi(G) \geq \omega(G)$.
\end{defn}

\begin{defn}
  A perfect graph is a graph that respects this condition for every induced subgraph:
  $$ \omega(G) = \chi(G)$$
\end{defn}

\begin{theorem}[Lovasz]
  $G$ is perfect if and only if $\overline{G}$ is perfect.
\end{theorem}

\subsection{Intersection graphs}

\begin{defn}
The \textit{intersection graph} of a collection $\zeta$ of objects is the graph
$(\zeta,E)$ such that $c_1c_2\in E \Leftrightarrow c_1 \cap c_2 \neq \varnothing$.
\end{defn}

An intersection can be seen as a relationship between two objects. In this thesis, it will be important to define these relations more formally to characterize intersection graphs.

\begin{defn}
  A partial order is a binary relation $\leq$ over a set $A$ satisfying these axioms:
  \begin{itemize}
    \item if $a \leq b$ and $b \leq a$ then $a = b$ (antisymmetry).
    \item if $a \leq b$ and $b \leq c$ then $a \leq c$ (transitivity).
    \item $a \leq a$ (reflexivity).
  \end{itemize}
\end{defn}

\begin{defn}
  A total order is a partial order where the reflexivity order is replaced by the connex property:
  $$a \leq b\ \text{or}\  b \leq a$$
\end{defn}

\begin{defn}
   A partially ordered set or poset  $(S,\leq)$ where $S$ a set and $\leq$ a partial
   order on $S$.
\end{defn}

\begin{defn}
  A spanning order $(V,<)$ of a graph $G = (V,E)$ is a total order on $V$ such that for any three vertices $u < v < w$:

  $$uw \in E \to uv \in E\ \text{or}\ vw \in E$$
\end{defn}


\begin{defn}
  A graph $G = (V,E)$ is a comparibility graph if there exists a partial order
  $\leq$ such that $vw \in E \Leftrightarrow v \leq w$ or $w \leq v$.
  Equivalently, $G$ is a comparability graph if it is the comparability graph of
  a poset. For example, the Hasse diagram (figure \ref{fig:hasse}) is a
  comparability graph where the relation is inclusion.
\end{defn}

\begin{defn}
  A graph $G = (V,E)$ is a co-comparability graph if its complement is a comparability graph.
\end{defn}

There are multiple characterizations for the co-comparability graph class; we will see one that uses a poset to characterize it:

\begin{theorem}[Damaschke \cite{damaschkeDistancesCocomparabilityGraphs1992}]
  \label{theo:spanning}
  A graph $G$ is a co-comparability graph if and only if it has a spanning order.
\end{theorem}


\begin{figure}
\centering

\begin{scaletikzpicturetowidth}{\textwidth}
\begin{tikzpicture}[scale=1.5]

  % poset
  \draw (-1cm,0cm) node (v2) {$\{x\}$};
  \draw (1cm,0cm)  node (v3) { $\{y\}$ };
  \draw (0cm,-1cm) node (v4) {$\varnothing$};
  \draw (0cm,1cm)  node (v1) {$\{x,y\}$};

  \draw[thick]  (v1) edge (v2);
  \draw[thick]  (v3) edge (v1);
  \draw[thick]  (v3) edge (v4);
  \draw[thick]  (v2) edge (v4);

  % graph
  \node[draw,circle,inner sep=2pt,fill,label distance=1cm] (v1g) at (3.5,1) {};
  \node[draw,circle,inner sep=2pt,fill,label distance=1cm] (v2g) at (3.5,-1) {};
  \node[draw,circle,inner sep=2pt,fill,label distance=1cm] (v3g) at (4.5,0) {};
  \node[draw,circle,inner sep=2pt,fill,label distance=1cm] (v4g) at (2.5,0) {};
  \draw  (v1g) edge (v2g);
  \draw  (v1g) edge (v4g);
  \draw  (v4g) edge (v2g);
  \draw  (v1g) edge (v3g);
  \draw  (v3g) edge (v2g);

\end{tikzpicture}
\end{scaletikzpicturetowidth}

\caption{On the left, Hasse diagram of a poset of the power set of 2 elements ordered by inclusion.
On the right, the comparability graph of this poset.}
\label{fig:hasse}
\end{figure}

\subsubsection{Disks}

A disk graph $G$ is a graph that is an intersection graph of disks on the plane, when the size
of the disk is unitary, we talk about unit disk graphs. This class of graphs
is important for this thesis, as thin strip graphs are a sub-class of
unit disk graphs.

We will refer to the unit disk graph class as UDG and an example of a realization
can be found in the figure \ref{fig:udg}.

\paragraph{Induced forbidden subgraphs} The characterization of this class with respect to
its induced forbidden subgraphs has been studied \cite{atminasForbiddenInducedSubgraphs2016}.

\begin{theorem}[Atminas-Zamaraev]
  For every integer $k > 1$, $\overline{K_2 + C_{2k+1}}$ is a minimal induced subgraph of UDG.
\end{theorem}

\begin{theorem}[Atminas-Zamaraev]
  For every integer $k > 4$, $\overline{C_{2k}}$ is a minimal induced subgraph of UDG.
\end{theorem}

\begin{figure}
\centering

\begin{scaletikzpicturetowidth}{\textwidth}
\begin{tikzpicture}[scale=1.5]

  \draw (-0.5,2.5) circle [radius=0.5];
  \draw (0,2.5) circle [radius=0.5];
  \draw (-0.5,2) circle [radius=0.5];
  \draw (0,2) circle [radius=0.5];


  \node[draw,circle,inner sep=2pt,fill,label distance=1cm] (v1) at (-2.5,2.5) {};

  \node[draw,circle,inner sep=2pt,fill,label distance=1cm] (v2) at (-2,2.2) {};
  \node[draw,circle,inner sep=2pt,fill,label distance=1cm] (v3) at (-2,2.8) {};
  \node[draw,circle,inner sep=2pt,fill,label distance=1cm] (v4) at (-1.5,2.5) {};

  \draw  (v3) edge (v2);
  \draw  (v4) edge (v1);
  \draw  (v3) edge (v1);
  \draw  (v4) edge (v2);
  \draw  (v3) edge (v4);
  \draw  (v1) edge (v2);

\end{tikzpicture}
\end{scaletikzpicturetowidth}

\caption{Realization of a UDG (Unit Disk Graph).}
\label{fig:udg}
\end{figure}
