\documentclass[12pt,a4paper,oneside]{book}


\usepackage{setspace}
\onehalfspacing

\makeindex
\usepackage{mathtools}
\usepackage{textcomp}
\usepackage{float}
\usepackage{fancyhdr}
\usepackage{makeidx}
\pagestyle{myheadings}
\fancyhf{}
\rhead[\leftmark]{thepage}
\usepackage[toc,page]{appendix}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{url}
\setlength {\marginparwidth}{2cm}
\usepackage{todonotes}

\usepackage[all]{xy}
\usepackage{graphicx}
\usepackage{units}
\usepackage{enumerate}
\usepackage[hidelinks]{hyperref}
\usepackage[]{algorithm2e} %algorithm


\usepackage{amssymb, amsmath, amsthm, pict2e}
 \pagestyle{myheadings}

\usepackage{pdfsync}
\usepackage{rotating}
\usepackage{multirow}
\usepackage[normalem]{ulem}
\usepackage{cancel}

\usepackage{color}
%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{pgf,tikz}
\usetikzlibrary{decorations,arrows, hobby, graphs, quotes}
\usetikzlibrary{shapes.misc, positioning}
%\usetikzlibrary{decorations.pathmorphing}
\usepgflibrary{decorations.pathreplacing}

\def\arr{\hbox{𝓇}}

\usepackage{environ}
\makeatletter
\newsavebox{\measure@tikzpicture}
\NewEnviron{scaletikzpicturetowidth}[1]{%
  \def\tikz@width{#1}%
  \def\tikzscale{1}\begin{lrbox}{\measure@tikzpicture}%
  \BODY
  \end{lrbox}%
  \pgfmathparse{#1/\wd\measure@tikzpicture}%
  \edef\tikzscale{\pgfmathresult}%
  \BODY
}

%%%%%%%%%%%%%%%%  start macros  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\let\emph\relax % there's no \RedeclareTextFontCommand
\DeclareTextFontCommand{\emph}{\bfseries\em}

\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{defn}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{question}[theorem]{Question}

\theoremstyle{plain}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{obs}[theorem]{Observation}
\newtheorem{prop}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}

\newtheorem{hyp}[theorem]{Hypothesis}
\newtheorem{alg}[theorem]{Algorithm}
\numberwithin{theorem}{section}
\newcommand{\powerset}{\raisebox{.15\baselineskip}{\Large\ensuremath{\wp}}}

%\newcommand{\qed}{\mbox{$\Box$}}
%\newcommand{\proof}{\medbreak\par\noindent{\bf Proof. }}
\newtheorem{proofpart}{Part}
\makeatletter
\@addtoreset{proofpart}{theorem}
\makeatother
 \newcommand{\GP}{{\vec{G}}_P}
\newcommand{\GN}{{\vec{G}}_N}

\newcommand{\cover}{\mathrel{\rlap{$\prec$}
                                \rlap{\hskip 0.7em $\cdot$}
                                 \phantom{\prec}}}

\newcommand{\re}{re}

\newcommand{\up}{\mbox{\rm{up}}}
\newcommand{\side}{\mbox{\rm{side}}}
\newcommand{\type}{\mbox{\rm{type}}}


\def\reals{{\mathbb R}}


\newcommand{\ahat}{{\hat{a}}}
\newcommand{\bhat}{{\hat{b}}}
\newcommand{\chat}{{\hat{c}}}
\newcommand{\dhat}{{\hat{d}}}
\newcommand{\bolda}{{\bf{a}}}
\newcommand{\boldb}{{\bf{b}}}
\newcommand{\boldc}{{\bf{c}}}
\newcommand{\boldd}{{\bf{d}}}

\newcommand{\iplus}{{\cal I}^+}
\newcommand{\iminus}{{\cal I}^-}
\newcommand{\ipm}{{\cal I}^{\pm}}
\newcommand{\ipmix}{{\cal I}^{{\cal M}}}

\newcommand{\upl}{{\cal U}^+}
\newcommand{\uminus}{{\cal U}^-}
\newcommand{\upm}{{\cal U}^{\pm}}
\newcommand{\upmix}{{\cal U}^{{\cal M}}}


\newcommand{\ppl}{{\cal P}^+}
\newcommand{\pminus}{{\cal P}^-}
\newcommand{\ppm}{{\cal P}^{\pm}}
\newcommand{\ppmix}{{\cal P}^{{\cal M}}}
\newcommand{\bpm}{{\cal B}^{\pm}}
\newcommand{\tpm}{{\cal T}^{\pm}}
\newcommand{\bpmix}{{\cal B}^{{\cal M}}}
\newcommand{\tpmix}{{\cal T}^{{\cal M}}}

%quoting

\definecolor{quotemark}{gray}{0.7}
\makeatletter
\newlength\origparskip

\newcommand{\fquote}{%
  \@ifnextchar[{\fquote@i}{\fquote@i[]}%]
}

\def\fquote@i[#1]{%
  \def\tempa{#1}%
  \@ifnextchar[{\fquote@ii}{\fquote@ii[]}%]
}%

\def\fquote@ii[#1]{%
  \def\tempb{#1}%
  %\vspace{1em}%
  \noindent%
  \begin{list}{}{%
    \setlength{\leftmargin}{0.3\textwidth}%
    \setlength{\rightmargin}{0.1\textwidth}%
    \setlength{\origparskip}{\parskip}%
  }%
    \item[]%
      \begin{picture}(0,0)%
        \put(-15,-8){\makebox(0,0){\scalebox{4}{%
          \textcolor{quotemark}{\textquotedblright}}}}%
      \end{picture}%
      \begingroup
      \itshape
      \ignorespaces}%

\def\endfquote{%
  \endgroup
  \par
  \raggedleft
  \ifx\tempa\empty
  \else
    {\bfseries --- \tempa\par}%
    \setlength{\parskip}{\origparskip}%
    \ifx\tempb\empty
    \else
      (\tempb)%
  \fi
  \par
  %\vspace{0.5em}%
  \end{list}}
\makeatother

\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

%%%%%%%%%%%%%%%%  end macros  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\parindent1em
%\parskip1.0ex

\begin{document}

\frontmatter
\begin{titlepage}
\begin{center}
\textbf{UNIVERSIT\'E LIBRE DE BRUXELLES}\\
\textbf{Facult\'e des Sciences}\\
\textbf{D\'epartement d'Informatique}
\vfill{}\vfill{}

{\Huge  Characterization and complexity of \vspace*{.5cm} \linebreak[4] Thin Strip Graphs}

% s \vspace*{.5cm}  \linebreak[4] possibly on many lines   \linebreak[4]  nevertheless a short title is better

{\Huge \par}
\begin{center}{\LARGE Abdeselam El-Haman Abdeselam}\end{center}{\Huge \par}
\vfill{}\vfill{}
\begin{flushright}{\large \textbf{Promotor :} Prof. Jean Cardinal}\\\hfill{}{\large Master Thesis in Computer Sciences}\\
{\large }\hfill{}{}\end{flushright}{\large\par}
\vfill{}\vfill{}\enlargethispage{3cm}
\textbf{Academic year 2018~-~2019}
\end{center}
\end{titlepage}
\newpage
\thispagestyle{empty}
\null

\newenvironment{vcenterpage}
{\newpage\thispagestyle{empty}
\vspace*{\fill}}
{\vspace*{\fill}\par\pagebreak}

\newpage
\thispagestyle{empty}
\vspace*{5cm}

\begin{fquote}[Cave Johnson][Portal 2]
  Science isn't about why -- it's about \emph{why not}.
\end{fquote}


\chapter*{Acknowledgment}
\thispagestyle{empty}

\noindent First of all, I would like to thank my thesis supervisor, Prof. Jean Cardinal for allowing me to be creative and by encouraging me to explore new areas of combinatorics. His courses on complexity and data structures showed me a new facet of computer science I fell in love with. I am also very grateful to him for showing me how to work rigorously during my thesis and also for encouraging me to continue my career in research.

I would also like to thank M. (now Dr.) Keno Merckx for all the help he has given me during the thesis. Keno has shared with me a lot of knowledge that has been useful for me. I will keep his advices preciously.

I also want to thank M. Gilles Geeraerts for all the projects that we shared together in the University. He is the first teacher I had at the University and since, he has been very helpful and open for every question I had. His contagious passion of teaching and pedagogy has pushed me forward to the same direction.

I thanks also the Algorithms group of ULB. Their courses and seminars have been precious for me, and their members have always been open to answer any question, either it was scientific or about my career.

I thank my partner, who has supported me through this whole period and encouraged me to never stop. I also thank my mother for not giving up on me when it wasn't that easy.

Last, but not least, I thank every classmate, colleague, teacher, first-year students, and every person I did not mention for giving me a great experience at the university that I will never forget.\\

Gracias, mamá.

\thispagestyle{empty}
\setcounter{page}{0}
\tableofcontents
\mainmatter
\setcounter{page}{1}

%% CHAPTERS %%
\input{intro/main}
\input{chap1/main}
\input{chap2/main}
\input{chap3/main}
\input{chap4/main}
\input{chap5/main}

%% CHAPTERS %%


\input{appendix/main}



\backmatter

\printindex % use makeindex to generate the index

\bibliographystyle{plain}

\bibliography{main} %use bibtex to generate the bibliography


\end{document}
